MSc Thesis

### 2023-04-30 manuscript/main.tex
more juli comments

We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application. We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis. We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it. After freeze drying, the samples for phenolic processing were ground under the same conditions as the other samples.

@@

To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}). We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet. We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon. After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples.

@@

@@

With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production. Over the past thirty years, research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}). Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}). Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\>0.75\% total dry weight) (\cite{reynolds_silicon_2016}). Silicon is highly abundant in many soils globally, and is the second most abundant element in the earth’s crust, behind only oxygen (\cite{ma_functions_2003}). It’s high expression in cereals, high abundance in many soils, and incredible broad spectrum effects on plant vigor and stress tolerance have make it a tantalizing target for improvements in agricultural yield and sustainability. Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.

Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress. For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005, katz_silicon_2019}). On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}). In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}). A current hypothesis explaining the broad-spectrum activity of silicon is presented in Coskun et al. (\citeyear{coskun_controversies_2019}), where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance [I could be more specific if needed]. Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts. Finding ways to improve crops towards increased silicon use efficiency is key to harnessing the benefits that plant silicon can confer.

>>>>

We harvested three fully expanded upper leaves the plants, and split the leaves in half along the midvein, 18 hours after treatment application. We placed one half of the tissue into a coin envelope, oven dried it for 4 days at 60\degree C, transferred it to a 2ml microcentrifuge tube with three 3.2 mm diameter steel beads, and ground it into a fine powder using a tisselyser (60 seconds at 30 Hz) in preparation for silicon content analysis. We placed the other half of the tissue into a microcentrifugre tube, flash froze it in liquid nitrogen, and subsequently freeze dried it. After freeze drying, we ground the samples for phenolic processing under the same conditions as the other samples.

@@

To measure the silicon content of the leaf tissue, we followed a modified version of the bench-top x-ray fluorescence (XRF) method (\cite{reidinger_rapid_2012}). We pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet. We then placed the pellet in an Olymus Vanta pXRF mounted in a benchtop stand, and used a 45 second scan time to quantify silicon. After each use, we cleaned the pellet die and XRF analyzer to minimize contamination between samples (comment left to include picture. Picture of the device or the pellet die?).

@@

We thank Dr. Aaron Beattie, Dr. Mazen Aljarrah, and Dr. Gurcharn Singh Brar for providing seeds for this experiment. Chelsea Gowton assisted with the harvesting and processing of samples. Dennis Chiu assisted with sample preparation and XRF analysis. Dr. Shaun Barker and [Dr. Brian McNulty] assisted with XRF analysis. Dr. Simone Castellarin and Dr. Joana Pico Carbajo advised and assisted with the development of the protocol for phenolic analysis.

@@

With a growing global population, and an increasingly imperiled biosphere, the quest for simultaneous increases in both the output and sustainability of agriculture has spurred development and research into new techniques that can help to feed the world and reduce the negative ecological impacts of large scale agricultural production. Over the past thirty years, increasing research momentum has gathered around plant silicon as a potential tool to effect sustainable increases in crop production, with particular applicability in the cereal crops (\cite{reynolds_silicon_2016,christian_breeding_2022}). Cereal crops are globally important, covering over one-third of the world’s arable land, making up over 50\% of the daily caloric intake for most people (\cite{faostat, rudel_agricultural_2009, awika_major_2011}). Cereals are members of the grass family (Poaceae) and typically have relatively high plant silicon content (\sim0.6 - 10\% total dry weight) (\cite{reynolds_silicon_2016}). Silicon's high abundance in many soils, high expression in cereal crops, and incredible broad spectrum effects on plant vigor and stress tolerance have made it a tantalizing target for improvements in agricultural yield and sustainability.

Silicon underpins a variety of physiological and developmental strategies that plants use to cope with stress. Though plants can complete their life cycle in the absence of silicon, its influence on such a diverse range of plant physiological functions has caused researchers to emphasize its importance relative to other non-essential nutrients.  For biotic stressors, silicon can reduce the damage plants experience from herbivory, increase resistance to fungal pathogens, and improve competitive ability with other organisms (\cite{fauteux_silicon_2005, katz_silicon_2019}). On the abiotic side, silicon supplementation improves plant resistance to soil salinity and heavy metal contamination, improves performance against temperature extremes and high irradiation, and helps plants to cope with drought stress (\cite{cooke_consistent_2016}). In comparing stressed plants grown in the absence or presence of silicon, Si+ plants showed a transcriptome profile similar to unstressed plants (\cite{coskun_controversies_2019}). A current hypothesis explaining the broad-spectrum activity of silicon is presented in Coskun et al. (\citeyear{coskun_controversies_2019}), where the authors suggest that silicon deposited in the apoplast of plant tissues where it modulates biological functions of the plant, and ecological interaction with natural enemies, yielding net positive increases in plant performance [I could be more specific if needed]. Realizing these beneficial effects depends on the plant’s ability to efficiently source silicon from the soil and uptake it in sufficient amounts. Finding ways to improve crops towards increased silicon use efficiency is key to harnessing the benefits that plant silicon can confer.
